Machine Learning, Architectural Styles and Property Values

### 2020-11-06 text/manuscript_assa.Rmd
Table updates

Will the resulting classifiers be biased with respect to property values, picking up correlated but irrelevant cues such as greenery, upkeep, or cars brands instead of building exteriors? Are, for instance, more valuable houses more likely to be classified as the popular *Late Victorian* style while starter homes are disproportionally more likely to be classified as e.g. *Early Victorian*? We expect that in cases in which the ensemble provided a high confidence score, building features are clearly detectable on the pictures. For close calls, however, where models within the ensemble cannot agree, other factors might tilt the scale towards an incorrect class. A sizable difference in estimated marginal prices for machine's classifications versus the estimates for architects' classifications in the subsequent hedonic regression (Eq. \ref{eq:hed}, below) would indicate such a bias.

@@

Residential real estate transactions are public data in the UK, collected and published by the Land Registry [@LandRegistry2016a]. The records include the date of transaction, price paid, street address, a classification of the property type (flat, detached, semi-detached, or terraced house), the estate type (freehold or leasehold) and an indicator for newly built properties. We select transactions from Cambridge proper which were recorded between January 1995 and October 2018, excluding any leaseholds, flats and properties classified as type "other", and sales with prices below £50,000 or in excess of £2,000,000. Linking recently taken images to sales data from 24 years assumes that houses have not changed their architectural styles -- which seems reasonable. We are not aware of any conversions of just the exterior of buildings. In case of full redevelopments, we exclude any sales for a given address pre-dating the redevelopment, ensuring that a currently taken pictures represents the house at all included transactions. Table \ref{tab:sumstats} presents summary statistics for the sample.

@@

Here, the natural logarithm of sales prices is explained by a linear combination of hedonic attributes described in vector $\mathbf{X}$, vectors of year $\mathbf{Y}$ and neighborhood $\bm{Loc}$ dummy variables and the building's estimated $\mathbf{Style}$ and the prevailing styles of other buildings in the direct proximity $(\mathbf{StyleNeigh})$.  $\mathbf{Style\cdot StyleNeigh}$ is a vector of interaction terms for the building's and the  neighborhood's dominant style. The intercept is denoted by $\alpha$ while $\bm{\beta}$, $\bm{\delta}$, $\bm{\eta}$, $\bm{\iota}$, $\bm{\gamma}$ and $\bm{\lambda}$ are vectors of regression coefficients. $\epsilon$ is the IID error term. Heteroscedastic robust standard errors will be reported.

Are buildings with different appearances imperfect substitutes catering to multiple groups of households with distinct style or vintage preferences? @Buitelaar2017 indicate that any premium for an architectural style must stem from either differences in construction prices (which they do not find in their Dutch sample) or from supply constraints, as new construction potentially does not capture the demand for traditional styles. For Cambridge, new supply will inevitably be of either contemporary or revival style as historic vintages (and their styles), by definition, are not supplied any more. Estimating Eq. \ref{eq:hed} for a subset of newly constructed buildings will show whether construction prices or supply constraints for new homes built according to different architectural styles persist -- if too few vernacular buildings were built, prices should reflect such a shortage. Singling out new buildings crucially controls for the otherwise unobserved age of buildings, which is tightly intertwined with its aesthetics.

@@

The majority of misclassification is concentrated in temporally adjacent eras: 10 percent of *Late Victorian/Edwardian* buildings are labelled as *Early Victorian* or 18 percent of *Postwar* buildings are erroneously regarded as stemming from the *Interwar* period. Hardly ever is a contemporary building mistaken for a historic home.

@@

The sample images classified by architects is not necessarily representative of the overall population so we re-estimate the model using machine classifications for *all* buildings for which we were able to extract usable pictures from Google Street View. Overall, the style coefficients become more positive relative to the base, implying that our automatic classification is biased and tends to classify unattractive buildings to the *Contemporary* style. Reversely, more attractive buildings are disproportionally often misclassified as *Late Vic./Edw.*, pushing up the corresponding coefficient to $0.11$. This bias vanishes in the third model, which uses only confidently estimated machine styles (Column 3).

Models 4--6 follow the same logic as 1--3 but are estimated for *new* buildings only. Basically this contrasts houses which might *look* different from the street, having either contemporary or revival facades, but which are all modern homes at their core. Differences in e.g. materials, floor plans, green space and gardens are minimal. For newly built homes, no effect of *Revival* architecture on price can be found for high confidence ML estimates, neither for classifications by machines nor by architects. After controlling for location, building characteristics and quality, buyers show no willingness to pay a premium for *Revival* architecture (Columns 4 and 6).

Introducing a variable capturing the most prominent style of buildings on the same street and within 100 m ($\mathbf{VintNeigh}$) does not change the pecking order of building styles much (Table \ref{tab:hedreg}, Column 7). Neighboring styles are compared to the base case of a *Contemporary* house in a *Contemporary* street. The negative coefficients on dissimilar neighboring styles support a finding by @Lindenthal2017 that shows that a harmonious match of a building's shape with its direct environment leads to a price premium -- or a discount in case of shape mismatches. In relative terms, *Postwar* buildings are the least popular neighbors. The difference between *Revival* and *Contemporary* neighbors is not statistically significant. We fail to find evidence in favor of positive externalities from revival architecture. Everything else equal, buildings within ensembles of *Contemporary* generally do not achieve higher transaction prices than buildings in more historic areas (as Table \ref{tab:comb} will confirm later).

@@

The large ground truth data set allows for a comparison of human expert versus machine classifications. Poor image quality, for instance obstructed views or the lack of informative features such as windows, is correlated with misclassifications. Importantly, for cases in which the ML classifiers are relatively indecisive between classes, visual cues that are correlated with the exterior's style but not part of a building's architecture (for instance trees or cars) bias the estimates. Fore these less certain cases, more valuable homes are systematically more likely to be classified as e.g. *Late Victorian/Edw.* and less likely to be considered *Contemporary*. A follow up study could investigate which elements in the building pictures lead to the misclassifications [@Ribeiro2016]. Given the danger of systematic biases, one should remain wary of ML estimates derived from data sets sizable enough to train models -- but too small to investigate any biases.

Third, we test whether revival or vernacular architecture leads to price premia over contemporary architecture. We do not find evidence for a preference of residents for specific architecture after controlling for quality and location of buildings. This applies to both direct price effects but also indirect effects of a buildings appearance on the value of neighboring homes. Further work will extend the spatial scope of the study area to the UK in general and to other countries.

@@

\label{tab:imgqualsumstats}
\begin{tabular}{lcccc} 
\toprule 
Architects' classifications & Share Blocked & House Area & Window Area & Image Offset \\ 
\midrule
 Georgian & 0.07 & 0.81 & 0.15 & 0.07 \\ 
 Early Vic. & 0.06 & 0.86 & 0.16 & 0.05 \\ 
 Late V./Edw. & 0.11 & 0.84 & 0.17 & 0.08 \\ 
 Interwar & 0.19 & 0.70 & 0.11 & 0.12 \\ 
 Postwar & 0.16 & 0.64 & 0.07 & 0.11 \\ 
 Cont. & 0.09 & 0.72 & 0.11 & 0.09 \\ 
 Revival & 0.10 & 0.78 & 0.11 & 0.09 \\ 
 \bottomrule
\end{tabular}
\begin{minipage}{0.9\textwidth}
\vspace{0.25cm}
\footnotesize \emph{Notes:} Using an Inception/Resnet object detection model trained on Open Images, basic object on the images are detected and the share of the image area taken up by cars, trees, buildings and windows are calculated. Images taken not taken at an optimal angle or zoom factor are detected by calculating the offset between the bounding box for the largest detected building and the center of the image.
\end{minipage}
\end{table}

@@

\end{tabular} 
 \begin{minipage}{\textwidth} 
\footnotesize 
\vspace{0.25cm}
\emph{Notes:} Cross-tabulation of photo characteristics by machine and architects' classifications. Off-diagonal elements ($Machine_{j}-Architects_{i}$) describe the difference in measure intensity between correctly and incorrectly classified photos for a given classification by architects. For example, cell \{Early Vic., Georgian\} compares the mean characteristic value for the set of photos which are actually Georgian but misclassified as Early Vic. to the set of correctly classified Georgian photos. The numbers in italic represent the mean characteristic value for each set of correctly classified photos. T-stats are in parenthesis.
\end{minipage} 
\endgroup 
\end{table}

@@

\begin{figure}[htb!]
  \caption{Difference of $F_1$-scores: spatial model vs. base model }
  \centering
    \includegraphics[width=0.8\textwidth]{figures/barplot_f1scores.jpg}
  \label{fig:boxplot}
\begin{minipage}{0.7\textwidth}
\vspace{0.25cm}
\footnotesize \emph{Notes:} The bar plot shows difference in the $F_1$-scores per category for the classifiers using both building level and neighbor information (dark and light blue) vs. a classifier using building level information only (red).
\end{minipage}
\end{figure}

@@

\begin{table}[htb!]
\caption{Machine style classifications as dependent variables (Eq. \ref{eq:rev} AIC comparison) }
\label{tab:aic}
\centering
\begin{tabular}{lrrrr}
  \toprule
 & \emph{No ML pred.} & \emph{Base pred.} & \emph{Spatial pred., all} & \emph{Spatial pred., high certainty} \\ 
  \midrule
  Georgian & 1,282 & 1,052 & 1,037 & 714 \\ 
  Early Victorian & 10,231 & 7,597 & 7,259 & 4,594 \\ 
  Late Vic./Edw. & 16,204 & 11,059 & 10,491 & 6,173 \\ 
  Interwar & 18,333 & 14,471 & 13,637 & 7,670 \\ 
  Postwar & 16,845 & 14,873 & 14,648 & 7,613 \\ 
  Contemporary & 2,885 & 2,281 & 2,240 & 1,054 \\ 
  Revival & 3,336 & 2,564 & 2,480 & 1,089 \\ 
   \bottomrule
\end{tabular}
\begin{minipage}{0.7\textwidth}
\vspace{0.25cm}
\footnotesize \emph{Notes:}
AIC values for 28 binomial models in which the occurrence of building styles (in rows) is explained by hedonic characteristics, location and time dummies (Column 1), augmented by the base ML predictor (Column 2), the suggested spatial ML predictor (Column 3 \& 4). Across building styles, the AIC are lowest for the spatial ML predictor, especially when uncertain predictions are omitted (Column 4). This suggests that the spatial predictor performs best.
\end{minipage}
\end{table}

>>>>

Will the resulting classifiers be biased with respect to property values, picking up correlated but irrelevant cues such as greenery, upkeep, or cars brands instead of building exteriors? Are, for instance, more valuable houses more likely to be classified as the popular *Late Victorian* style while starter homes are disproportionally more likely to be classified as e.g. *Early Victorian*? We expect that in cases in which the ensemble provided a high confidence score, building features are clearly detectable on the pictures. For close calls, however, where models within the ensemble cannot agree, other factors might tilt the scale towards an incorrect class. A sizable difference in estimated marginal prices for machine classifications versus the estimates for architects' classifications in the subsequent hedonic regression (Eq. \ref{eq:hed}, below) would indicate such a bias.

@@

Residential real estate transactions are public data in the UK, collected and published by the Land Registry [@LandRegistry2016a]. The records include the date of transaction, price paid, street address, a classification of the property type (flat, detached, semi-detached, or terraced house), the estate type (freehold or leasehold) and an indicator for newly built properties. We select transactions from Cambridge which were recorded between January 1995 and October 2018, excluding any leaseholds, flats and properties classified as type "other", and sales with prices below £50,000 or in excess of £2,000,000. Linking recently taken images to sales data from 24 years assumes that houses have not changed their architectural styles -- which seems reasonable. We are not aware of any conversions of just the exterior of buildings. In case of full redevelopments, we exclude any sales for a given address pre-dating the redevelopment, ensuring that a currently taken picture represents the house at all included transactions. Table \ref{tab:sumstats} presents summary statistics for the sample.

@@

Here, the natural logarithm of sales prices is explained by a linear combination of hedonic attributes described in vector $\mathbf{X}$, vectors of year $\mathbf{Y}$ and neighborhood $\bm{Loc}$ dummy variables and the building's estimated $\mathbf{Style}$ and the prevailing styles of other buildings in the direct proximity $(\mathbf{StyleNeigh})$.  $\mathbf{Style\cdot StyleNeigh}$ is a vector of interaction terms for the building's and the  neighborhood's dominant style. The intercept is denoted by $\alpha$ while $\bm{\beta}$, $\bm{\delta}$, $\bm{\eta}$, $\bm{\iota}$, $\bm{\gamma}$ and $\bm{\lambda}$ are vectors of regression coefficients. $\epsilon$ is the IID error term. Heteroscedasticity robust standard errors will be reported.

Are buildings with different appearances imperfect substitutes catering to multiple groups of households with distinct style or vintage preferences? @Buitelaar2017 indicate that any premium for an architectural style must stem from either differences in construction prices (which they do not find in their Dutch sample) or from supply constraints, as new construction potentially does not capture the demand for traditional styles. For Cambridge, new supply will inevitably be of either contemporary or revival style as historic vintages (and their styles), by definition, are not supplied any more. Estimating Eq. \ref{eq:hed} for a subset of newly constructed buildings will show whether construction prices or supply constraints for new homes built according to different architectural styles prevail -- if too few vernacular buildings were built, prices should reflect such a shortage. Singling out new buildings crucially controls for the otherwise unobserved age of these buildings, which is tightly intertwined with their aesthetics.

@@

The majority of misclassifications is concentrated in temporally adjacent eras: 10 percent of *Late Victorian/Edwardian* buildings are labelled as *Early Victorian* or 18 percent of *Postwar* buildings are erroneously regarded as stemming from the *Interwar* period. Hardly ever is a contemporary building mistaken for a historic home.

@@

The sample of images classified by architects is not necessarily representative of the overall population so we re-estimate the model using machine classifications for *all* buildings for which we were able to extract usable pictures from Google Street View. Overall, the style coefficients become more positive relative to the base, implying that our automatic classification is biased and tends to classify bland or low quality buildings to the *Contemporary* style. Reversely, more attractive or higher quality buildings are disproportionately often misclassified as *Late Vic./Edw.*, pushing up the corresponding coefficient to $0.11$. This bias vanishes in the third model, which uses only confidently estimated machine styles (Column 3).

Models 4--6 follow the same logic as 1--3 but are estimated for *new* buildings only, which are either in the contemporary or the revival style. Basically this contrasts houses which might *look* different from the street, having either contemporary or revival facades, but which are all modern homes at their core. Differences in e.g. materials, floor plans, green space and gardens are minimal. For newly built homes, no effect of *Revival* architecture on price can be found for high confidence ML estimates, neither for classifications by machines nor by architects. After controlling for location, building characteristics and quality, buyers show no willingness to pay a premium for *Revival* architecture (Columns 4 and 6).

Introducing a variable capturing the most prominent style of buildings on the same street and within 100 m ($\mathbf{StyleNeigh}$) does not change the pecking order of building styles much (Table \ref{tab:hedreg}, Column 7). Neighboring styles are compared to the base case of a *Contemporary* house in a *Contemporary* street. The negative coefficients on dissimilar neighboring styles support a finding by @Lindenthal2017 that shows that a harmonious match of a building's shape with its direct environment leads to a price premium -- or a discount in case of shape mismatches. In relative terms, *Postwar* buildings are the least popular neighbors. The difference between *Revival* and *Contemporary* neighbors is not statistically significant. We fail to find evidence in favor of positive externalities from revival architecture. Everything else equal, buildings within ensembles of *Contemporary* generally do not achieve higher transaction prices than buildings in more historic areas (as Table \ref{tab:comb} will confirm later).

@@

The large ground truth data set allows for a comparison of human expert versus machine classifications. Poor image quality, for instance obstructed views or the lack of informative features such as windows, is correlated with misclassifications. Importantly, for cases in which the ML classifiers are relatively indecisive between classes, visual cues that are correlated with the exterior's style but not part of a building's architecture (for instance trees or cars) bias the estimates. For these less certain cases, more valuable homes are systematically more likely to be classified as e.g. *Late Victorian/Edw.* and less likely to be considered *Contemporary*. A follow up study could investigate which elements in the building pictures lead to the misclassifications [@Ribeiro2016]. Given the danger of systematic biases, one should remain wary of ML estimates derived from data sets sizable enough to train models -- but too small to investigate any biases.

Third, we test whether revival or vernacular architecture leads to price premia over contemporary architecture. We do not find evidence for a preference of residents for specific architecture after controlling for quality and location of buildings. This applies to both direct price effects but also indirect effects of a building's appearance on the value of neighboring homes. Further work will extend the spatial scope of the study area to the UK in general and to other countries.

@@

\label{tab:imgqualsumstats}
\begin{tabular}{lcccc} 
\toprule 
Architects' classifications & Share Blocked & House Area & Window Area & Image Offset \\ 
\midrule
 Georgian & 0.07 & 0.81 & 0.15 & 0.07 \\ 
 Early Vic. & 0.06 & 0.86 & 0.16 & 0.05 \\ 
 Late V./Edw. & 0.11 & 0.84 & 0.17 & 0.08 \\ 
 Interwar & 0.19 & 0.70 & 0.11 & 0.12 \\ 
 Postwar & 0.16 & 0.64 & 0.07 & 0.11 \\ 
 Cont. & 0.09 & 0.72 & 0.11 & 0.09 \\ 
 Revival & 0.10 & 0.78 & 0.11 & 0.09 \\ 
 \bottomrule
\end{tabular}
\begin{minipage}{0.9\textwidth}
\vspace{0.25cm}
\footnotesize \emph{Notes:} Using an Inception/Resnet object detection model trained on Open Images, basic object on the images are detected and the share of the image area taken up by cars, trees, buildings and windows are calculated. Images not taken at an optimal angle or zoom factor are detected by calculating the offset between the bounding box for the largest detected building and the center of the image.
\end{minipage}
\end{table}

@@

\end{tabular} 
 \begin{minipage}{\textwidth} 
\footnotesize 
\vspace{0.25cm}
\emph{Notes:} Cross-tabulation of photo characteristics by machine and architects' classifications. Off-diagonal elements ($Machine_{j}-Architects_{i}$) describe the difference in measure intensity between correctly and incorrectly classified photos for a given classification by architects. For example, cell \{Early Vic., Georgian\} compares the mean characteristic value for the set of photos which are actually Georgian but misclassified as Early Vic. to the set of correctly classified Georgian photos. T-stats are in parenthesis.
\end{minipage} 
\endgroup 
\end{table}

@@

\begin{figure}[htb!]
  \caption{Difference of $F_1$-scores: spatial model vs. base model }
  \centering
    \includegraphics[width=0.8\textwidth]{figures/barplot_f1scores.jpg}
  \label{fig:boxplot}
\begin{minipage}{0.7\textwidth}
\vspace{0.25cm}
\footnotesize \emph{Notes:} The bar plot shows differences in the $F_1$-scores per category for the classifiers using both building level and neighbor information (dark and light blue) vs. a classifier using building level information only (red).
\end{minipage}
\end{figure}

@@

\begin{table}[htb!]
\caption{Machine style classifications as dependent variables (Eq. \ref{eq:rev} AIC comparison) }
\label{tab:aic}
\centering
\begin{tabular}{lrrrr}
  \toprule
 & \emph{No ML pred.} & \emph{Base pred.} & \emph{Spatial pred., all} & \emph{Spatial pred., high certainty} \\ 
  \midrule
  Georgian & 1,282 & 1,052 & 1,037 & 714 \\ 
  Early Victorian & 10,231 & 7,597 & 7,259 & 4,594 \\ 
  Late Vic./Edw. & 16,204 & 11,059 & 10,491 & 6,173 \\ 
  Interwar & 18,333 & 14,471 & 13,637 & 7,670 \\ 
  Postwar & 16,845 & 14,873 & 14,648 & 7,613 \\ 
  Contemporary & 2,885 & 2,281 & 2,240 & 1,054 \\ 
  Revival & 3,336 & 2,564 & 2,480 & 1,089 \\ 
   \bottomrule
\end{tabular}
\begin{minipage}{0.7\textwidth}
\vspace{0.25cm}
\footnotesize \emph{Notes:}
AIC values for 28 binomial models in which the occurrence of building styles (in rows) is explained by hedonic characteristics, location and time dummies (Column 1), augmented by the base ML predictor (Column 2), and the suggested spatial ML predictor (Column 3 \& 4). Across building styles, the AIC are lowest for the spatial ML predictor, especially when uncertain predictions are omitted (Column 4). This suggests that the spatial predictor performs best.
\end{minipage}
\end{table}

### 2020-11-06 text/manuscript_assa.tex
Table updates

  \title{Machine Learning, Architectural Styles and Property Values}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson
(University of Alabama)}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{13 April 2020}

@@

Residential real estate transactions are public data in the UK,
collected and published by the Land Registry (Land Registry 2017). The
records include the date of transaction, price paid, street address, a
classification of the property type (flat, detached, semi-detached, or
terraced house), the estate type (freehold or leasehold) and an
indicator for newly built properties. We select transactions from
Cambridge proper which were recorded between January 1995 and October
2018, excluding any leaseholds, flats and properties classified as type
``other'', and sales with prices below £50,000 or in excess of
£2,000,000. Linking recently taken images to sales data from 24 years
assumes that houses have not changed their architectural styles -- which
seems reasonable. We are not aware of any conversions of just the
exterior of buildings. In case of full redevelopments, we exclude any
sales for a given address pre-dating the redevelopment, ensuring that a
currently taken pictures represents the house at all included
transactions. Table \ref{tab:sumstats} presents summary statistics for
the sample.

@@

Here, the natural logarithm of sales prices is explained by a linear
combination of hedonic attributes described in vector \(\mathbf{X}\),
vectors of year \(\mathbf{Y}\) and neighborhood \(\bm{Loc}\) dummy
variables and the building's estimated \(\mathbf{Style}\) and the
prevailing styles of other buildings in the direct proximity
\((\mathbf{StyleNeigh})\). \(\mathbf{Style\cdot StyleNeigh}\) is a
vector of interaction terms for the building's and the neighborhood's
dominant style. The intercept is denoted by \(\alpha\) while
\(\bm{\beta}\), \(\bm{\delta}\), \(\bm{\eta}\), \(\bm{\iota}\),
\(\bm{\gamma}\) and \(\bm{\lambda}\) are vectors of regression
coefficients. \(\epsilon\) is the IID error term. Heteroscedastic robust
standard errors will be reported.

Are buildings with different appearances imperfect substitutes catering
to multiple groups of households with distinct style or vintage
preferences? Buitelaar and Schilder (2017) indicate that any premium for
an architectural style must stem from either differences in construction
prices (which they do not find in their Dutch sample) or from supply
constraints, as new construction potentially does not capture the demand
for traditional styles. For Cambridge, new supply will inevitably be of
either contemporary or revival style as historic vintages (and their
styles), by definition, are not supplied any more. Estimating Eq.
\ref{eq:hed} for a subset of newly constructed buildings will show
whether construction prices or supply constraints for new homes built
according to different architectural styles persist -- if too few
vernacular buildings were built, prices should reflect such a shortage.
Singling out new buildings crucially controls for the otherwise
unobserved age of buildings, which is tightly intertwined with its
aesthetics.

@@

The sample images classified by architects is not necessarily
representative of the overall population so we re-estimate the model
using machine classifications for \emph{all} buildings for which we were
able to extract usable pictures from Google Street View. Overall, the
style coefficients become more positive relative to the base, implying
that our automatic classification is biased and tends to classify
unattractive buildings to the \emph{Contemporary} style. Reversely, more
attractive buildings are disproportionally often misclassified as
\emph{Late Vic./Edw.}, pushing up the corresponding coefficient to
\(0.11\). This bias vanishes in the third model, which uses only
confidently estimated machine styles (Column 3).

Models 4--6 follow the same logic as 1--3 but are estimated for
\emph{new} buildings only. Basically this contrasts houses which might
\emph{look} different from the street, having either contemporary or
revival facades, but which are all modern homes at their core.
Differences in e.g.~materials, floor plans, green space and gardens are
minimal. For newly built homes, no effect of \emph{Revival} architecture
on price can be found for high confidence ML estimates, neither for
classifications by machines nor by architects. After controlling for
location, building characteristics and quality, buyers show no
willingness to pay a premium for \emph{Revival} architecture (Columns 4
and 6).

Introducing a variable capturing the most prominent style of buildings
on the same street and within 100 m (\(\mathbf{VintNeigh}\)) does not
change the pecking order of building styles much (Table
\ref{tab:hedreg}, Column 7). Neighboring styles are compared to the base
case of a \emph{Contemporary} house in a \emph{Contemporary} street. The
negative coefficients on dissimilar neighboring styles support a finding
by Lindenthal (2017b) that shows that a harmonious match of a building's
shape with its direct environment leads to a price premium -- or a
discount in case of shape mismatches. In relative terms, \emph{Postwar}
buildings are the least popular neighbors. The difference between
\emph{Revival} and \emph{Contemporary} neighbors is not statistically
significant. We fail to find evidence in favor of positive externalities
from revival architecture. Everything else equal, buildings within
ensembles of \emph{Contemporary} generally do not achieve higher
transaction prices than buildings in more historic areas (as Table
\ref{tab:comb} will confirm later).

@@

The large ground truth data set allows for a comparison of human expert
versus machine classifications. Poor image quality, for instance
obstructed views or the lack of informative features such as windows, is
correlated with misclassifications. Importantly, for cases in which the
ML classifiers are relatively indecisive between classes, visual cues
that are correlated with the exterior's style but not part of a
building's architecture (for instance trees or cars) bias the estimates.
Fore these less certain cases, more valuable homes are systematically
more likely to be classified as e.g.~\emph{Late Victorian/Edw.} and less
likely to be considered \emph{Contemporary}. A follow up study could
investigate which elements in the building pictures lead to the
misclassifications (Ribeiro, Singh, and Guestrin 2016). Given the danger
of systematic biases, one should remain wary of ML estimates derived
from data sets sizable enough to train models -- but too small to
investigate any biases.

Third, we test whether revival or vernacular architecture leads to price
premia over contemporary architecture. We do not find evidence for a
preference of residents for specific architecture after controlling for
quality and location of buildings. This applies to both direct price
effects but also indirect effects of a buildings appearance on the value
of neighboring homes. Further work will extend the spatial scope of the
study area to the UK in general and to other countries.

>>>>

  \title{Machine Learning, Architectural Styles and Property Values}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson
(University of Alabama)}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{14 April 2020}

@@

Residential real estate transactions are public data in the UK,
collected and published by the Land Registry (Land Registry 2017). The
records include the date of transaction, price paid, street address, a
classification of the property type (flat, detached, semi-detached, or
terraced house), the estate type (freehold or leasehold) and an
indicator for newly built properties. We select transactions from
Cambridge which were recorded between January 1995 and October 2018,
excluding any leaseholds, flats and properties classified as type
``other'', and sales with prices below £50,000 or in excess of
£2,000,000. Linking recently taken images to sales data from 24 years
assumes that houses have not changed their architectural styles -- which
seems reasonable. We are not aware of any conversions of just the
exterior of buildings. In case of full redevelopments, we exclude any
sales for a given address pre-dating the redevelopment, ensuring that a
currently taken picture represents the house at all included
transactions. Table \ref{tab:sumstats} presents summary statistics for
the sample.

@@

Here, the natural logarithm of sales prices is explained by a linear
combination of hedonic attributes described in vector \(\mathbf{X}\),
vectors of year \(\mathbf{Y}\) and neighborhood \(\bm{Loc}\) dummy
variables and the building's estimated \(\mathbf{Style}\) and the
prevailing styles of other buildings in the direct proximity
\((\mathbf{StyleNeigh})\). \(\mathbf{Style\cdot StyleNeigh}\) is a
vector of interaction terms for the building's and the neighborhood's
dominant style. The intercept is denoted by \(\alpha\) while
\(\bm{\beta}\), \(\bm{\delta}\), \(\bm{\eta}\), \(\bm{\iota}\),
\(\bm{\gamma}\) and \(\bm{\lambda}\) are vectors of regression
coefficients. \(\epsilon\) is the IID error term. Heteroscedasticity
robust standard errors will be reported.

Are buildings with different appearances imperfect substitutes catering
to multiple groups of households with distinct style or vintage
preferences? Buitelaar and Schilder (2017) indicate that any premium for
an architectural style must stem from either differences in construction
prices (which they do not find in their Dutch sample) or from supply
constraints, as new construction potentially does not capture the demand
for traditional styles. For Cambridge, new supply will inevitably be of
either contemporary or revival style as historic vintages (and their
styles), by definition, are not supplied any more. Estimating Eq.
\ref{eq:hed} for a subset of newly constructed buildings will show
whether construction prices or supply constraints for new homes built
according to different architectural styles prevail -- if too few
vernacular buildings were built, prices should reflect such a shortage.
Singling out new buildings crucially controls for the otherwise
unobserved age of these buildings, which is tightly intertwined with
their aesthetics.

@@

The sample of images classified by architects is not necessarily
representative of the overall population so we re-estimate the model
using machine classifications for \emph{all} buildings for which we were
able to extract usable pictures from Google Street View. Overall, the
style coefficients become more positive relative to the base, implying
that our automatic classification is biased and tends to classify bland
or low quality buildings to the \emph{Contemporary} style. Reversely,
more attractive or higher quality buildings are disproportionately often
misclassified as \emph{Late Vic./Edw.}, pushing up the corresponding
coefficient to \(0.11\). This bias vanishes in the third model, which
uses only confidently estimated machine styles (Column 3).

Models 4--6 follow the same logic as 1--3 but are estimated for
\emph{new} buildings only, which are either in the contemporary or the
revival style. Basically this contrasts houses which might \emph{look}
different from the street, having either contemporary or revival
facades, but which are all modern homes at their core. Differences in
e.g.~materials, floor plans, green space and gardens are minimal. For
newly built homes, no effect of \emph{Revival} architecture on price can
be found for high confidence ML estimates, neither for classifications
by machines nor by architects. After controlling for location, building
characteristics and quality, buyers show no willingness to pay a premium
for \emph{Revival} architecture (Columns 4 and 6).

Introducing a variable capturing the most prominent style of buildings
on the same street and within 100 m (\(\mathbf{StyleNeigh}\)) does not
change the pecking order of building styles much (Table
\ref{tab:hedreg}, Column 7). Neighboring styles are compared to the base
case of a \emph{Contemporary} house in a \emph{Contemporary} street. The
negative coefficients on dissimilar neighboring styles support a finding
by Lindenthal (2017b) that shows that a harmonious match of a building's
shape with its direct environment leads to a price premium -- or a
discount in case of shape mismatches. In relative terms, \emph{Postwar}
buildings are the least popular neighbors. The difference between
\emph{Revival} and \emph{Contemporary} neighbors is not statistically
significant. We fail to find evidence in favor of positive externalities
from revival architecture. Everything else equal, buildings within
ensembles of \emph{Contemporary} generally do not achieve higher
transaction prices than buildings in more historic areas (as Table
\ref{tab:comb} will confirm later).

@@

The large ground truth data set allows for a comparison of human expert
versus machine classifications. Poor image quality, for instance
obstructed views or the lack of informative features such as windows, is
correlated with misclassifications. Importantly, for cases in which the
ML classifiers are relatively indecisive between classes, visual cues
that are correlated with the exterior's style but not part of a
building's architecture (for instance trees or cars) bias the estimates.
For these less certain cases, more valuable homes are systematically
more likely to be classified as e.g.~\emph{Late Victorian/Edw.} and less
likely to be considered \emph{Contemporary}. A follow up study could
investigate which elements in the building pictures lead to the
misclassifications (Ribeiro, Singh, and Guestrin 2016). Given the danger
of systematic biases, one should remain wary of ML estimates derived
from data sets sizable enough to train models -- but too small to
investigate any biases.

Third, we test whether revival or vernacular architecture leads to price
premia over contemporary architecture. We do not find evidence for a
preference of residents for specific architecture after controlling for
quality and location of buildings. This applies to both direct price
effects but also indirect effects of a building's appearance on the
value of neighboring homes. Further work will extend the spatial scope
of the study area to the UK in general and to other countries.
